Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-06-16 analysis/memo2_TAS_reviewers.Rmd
revised "cog class" discussion

This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and a reviewer.

@@

By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias. For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors". If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."

We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", for two reasons. First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury. Second, at the time any strike occurs, not every prospective juror could have been struck. As we assume in the paper (see pages XXX), only a subset of prospective jurors present during jury selection are ever subject to strike at any one time. The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.

@@

Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$). As we discuss in the paper (page X, lines YY-YY), it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility. We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases. Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race). However, in principle, it could be modified to add that kind of feature.

>>>>

This memorandum replies to comments received in May 2023 from the TAS editor, a TAS associate editor, and an anonymous reviewer.

@@

We now introduce the term "cognizable class" later, in section 2. By "cognizable" class, we mean the social category to which the struck juror belongs that is the possible basis of illegal strike bias. For example, if the focus is on strike bias against Black prospective jurors because of their race, the "cognizable" class is "Black prospective jurors". If, however, the focus is on strike bias against Black female prospective jurors because of their race and gender, the "cognizable" class is "Black female prospective jurors."  Because _Batson_ applies without regard to the "majority" or "marginalized" status of struck jurors (e.g., Black prospective jurors _and_ White prospective jurors), we have avoided those terms.

We use the phrase "cognizable class members subject to strike" instead of "cognizable class jurors", for two reasons. First, at the time any strike occurs, no one is a "juror", because no one has yet been seated onto the jury. Second, at the time any strike occurs, not every prospective juror could have been struck. As we assume in the paper, only a subset of prospective jurors present during jury selection are ever subject to strike at any one time. The size of that subset is usually based on the number of seats of the jury itself plus the total number of strikes that the parties could exercise.

@@

Subsetting the historical data does not necessarily reduce the degree of incompatibility between the historical strike data and the current trial ($b_hist$ - $b_curr$). As we discuss in the paper, it depends on what one assumes about how much the subsetting criteria (e.g., the salient criteria for "similar" cases) changes the degree of incompatibility. We stress "assumes", because the true values of $b_hist$ and $b_curr$ are not directly observable in actual cases. Accordingly, for any subset of the historical strike data one chooses, our guidance for choosing an $\alpha$ value remains the same: Check for how sensitive the posterior estimate of illegal strike bias is to different values of $\alpha$. The current version of the software prototype does not let users subset the historical data by some characteristic (e.g., date, defendant race). However, in principle, it could be modified to add that kind of feature.
